NRM-tool (Numerical Reproducibility Measurement): a debugging tool to identify numerical errors and assess reproducibility of pipelines

### 2019-04-11 paper.tex
re-write in GigaScience journal format

\documentclass{article}

\usepackage{pdfcomment}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{booktabs}
\usepackage{makecell}

\usepackage[flushleft]{threeparttable}

\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}

\newcommand{\reprozip}[0]{ReproZip}

\title{Reprotool: a debugging tool to identify numerical errors 
and assess reproducibility of pipelines}

\author{M. Ali Salari, Lalet Scaria, Gregory Kiar, Lindsay B. Lewis,
  Alan C. Evans, Tristan Glatard}

@@

@@

\abstract{

Many experiments show that computational analysis results are still not 
completely reproducible. In particular, computational environments 
including different operating systems, hardware categories, and 
software versions are known to have effects on the results produced by 
analysis pipelines. These effects are presumably due to the creation, 
propagation and amplification of small numerical errors across the 
pipelines. 
Although new techniques such as virtualization , version 
control systems , and provenance management tools provide more reliable 
environment and significantly improve the reproducibility of scientific 
findings. 
However, the precise causes of such instabilities and the path along 
which they propagate in the pipelines are unclear.  We present a 
technique to identify the processes in the pipeline that create 
numerical errors along the execution, and we apply this technique to 
the HCP structural pre-processing pipelines.

@@

Research findings are expected to be reproducible so that 
their authenticity and reliability can be evaluated. 
There are different definitions for the terms reproducibility, 
and replicability, which leads to confusion since the 
same words are used for different concepts.

According to Peng's definition~\cite{peng2011reproducible}, 
reproducibility is defined as the ability to regenerate the same 
results as the original findings when the experiment is reanalyzed 
given exactly the same analytic methods and data. Reproducibility 
ensures that independent scientists can reproduce the same results 
using the same data and procedure as published in the original 
publication. On the contrary, replicability is defined as the ability 
to obtain similar results as published in the original study when 
the experiment is reimplemented using independent data and analytic 
methods. Replicability confirms scientific claims and ensures that 
independent investigators can produce consistent results, using new 
data and methods.

In addition, numerical reproducibility is defined as the ability to 
regenerate bit for bit identical results from multiple 
runs~\cite{hill2017numerical}. Two files will be considered numerically 
reproducible if they have identical binary contents. Binary 
comparison is calculated by comparing checksums. It must be pointed out 
that a computation might be reproducible based on Peng's definition, 
but not be numerically reproducible. For instance, small numerical 
errors created during the pipeline execution may hamper numerical 
reproducibility, but be negligible in the final results.

Reproducibility, as the cornerstone of scientific research, guarantees 
the reliability of results. A reproducible study provides a context in 
which one is able to get results that are consistent with the original 
work. In addition, it not only saves a great deal of time, but also 
enables others to use existing works as a part of their 
experiments~\cite{plesser2018reproducibility}. 
In this paper, we follow Peng's definition of reproducibility 
unless we directly refer to numerical reproducibility. 
We seek to identify the reasons why such reproducibility may not be 
ensured, focusing in particular on computational aspects.

Recently, scientists began to realize that the results of many 
scientific experiments were neither replicable nor reproducible. 
Results show that the variety of computational infrastructures 
including workstation types, parallelization methods, operating 
systems, and analysis packages are known to influence reproducibility 
because of the creation of small numerical 
errors~\cite{Gronenschild2012, diethelm2012limits, Glatard2015}. The 
propagation and amplification of these tiny errors by analysis 
pipelines may cause reproducibility issues. 
\note{In this case, the analysis pipelines are said to be numerically 
unstable. Numerical instability is a characteristic of the pipelines 
which amplify small numerical errors and then hamper the 
reproducibility of the analyses depending on the length of the pipeline 
and magnitude of the errors. In many cases, numerical instability is an 
important issue for reproducibility. }
In particular, the effect of operating system is quantified 
in~\cite{Glatard2015, Gronenschild2012} more specifically on some of 
the main neuroimaging pipelines including FSL, CIVET and Freesurfer 
pipelines. The results of comparisons indicate a significant errors 
between analysis results of the same experiment on different operating 
system. It should be noted that \emph{error} is defined as the binary 
differences between results of two analyses in this paper, which is 
calculated by the checksum method.

\note{The operating system is not the only part of the computational 
environment that may hamper reproducibility. Many other computational 
factors including hardware configuration, software version, and 
parallelization techniques are reported as 
the main influential computing infrastructure on reproducibility. }

There are different solutions to make reproducible analysis including 
containerization techniques to encapsulate software/hardware 
dependencies, provenance capturing tools, and version control systems. 
However, a comprehensive solution requires to fix the numerical 
instabilities. For this purpose, we introduce a framework to identify 
the processes in pipeline that create numerical errors across different 
runs. In this paper, the execution results of pipelines across 
different operating systems are considered to find out the cause of the 
irreproducibility using an interposition techniques.

Figure~\ref{fig:overview} illustrates an overview of the proposed 
technique. First, we execute pipelines in a consistence environment 
using docker container and Boutiques framework. In the next step, the 
processes that create differences are classified using an iterative 
approach. Finally, we make corresponding process graph of the 
pipeline execution that shows the processes related to the creation of 
errors. We will explain each step in detail in the following sections.

\begin{figure}[H]
\centering
  \includegraphics[scale=0.5]{images/overview.png}
  \caption{An overview of the proposed technique.}
  \label{fig:overview}
\end{figure}

The reminder of this paper explains the Reprotool in different parts as 
the execution environment of the pipelines, methods to measure 
differences, and how to characterize errors. We describe the HCP 
preprocessing pipelines as a popular neuroimaging pipelines to test and 
apply the Reprotool on. Finally, we report the processes that are 
responsible for the errors in the pipelines.

\section{Processing Environment}

We build the containerized version of pipelines for each operating 
system using docker technology which encapsulated whole the pipeline 
dependencies. The docker images consists of an specific computing 
environment such as operating system version, hardware configuration, 
software version, etc. Furthermore, derived images can be stored on 
the dockerhub repositories and be version controlled in the same way as 
the analysis code so that the exact same environment can be used to run 
an experiment.

To simplify computational reproducibility of the analysis, we execute 
docker images using Boutiques framework based on the command-line 
descriptions~\cite{glatard2017boutiques}. The Boutiques descriptor 
specifies the parameters and implementation of the experiment, and the 
invocation schema describes the parameter values such as input/output 
data path.

We processed subjects twice in every condition to identify possible 
\emph{intra-condition} errors coming from the use of pseudo-random 
numbers or other external factors. To further ensure that the observed 
errors are caused by operating system updates (\emph{inter-os 
differences}) and not by other factors, we implemented the following 
sanity checks:

\begin{itemize}

\item The file checksums of every subject are computed before and after 
processing to detect potential corruption during file transfers and 
other manipulations. 
\item The checksum of the Docker container images is recorded with
  every execution to identify any unintentional package update. In 
  addition, the list of packages with versions is printed by every task 
  and checked for consistency.
\item Hardware information is captured to make sure that errors are
  not coming from different CPU architectures.
\item Every task is executed on its own copy of the data to ensure
  that input data is not modified by the pipeline. 
\end{itemize}

\section{Measuring Differences}

We compute a file difference matrix between files produced for every 
subject in every operating system based on MD5 
checksums~\cite{md51992url} as in~\cite{Scaria2017}. MD5 is suitable to 
identify incidental file differences, although it is not 
cryptographically secure. Files produced only in some subjects or in 
some operating systems were left out.

In addition, we are able to quantify errors in the image files using 
the Dice coefficient and normalized root mean square error (NRMSE) 
computed as follows:

\begin{center}
\begin{equation}
RMSE = {\sqrt {\frac{1} {n}{\sum\limits_{t = 1}^n {(\hat{y}_{t} - {y}_{t} } })^{2} } }
\end{equation}
\end{center}

\begin{center}
\begin{equation}
NRMSE = {\frac{RMSE} {y_{max} - y_{min}}}
\end{equation}
\end{center}

\section{Pipeline Analysis: Characterizing Errors}

Once errors between conditions are quantified, we automatically
identify the steps in the pipeline responsible for such differences.
We use the \reprozip tool~\cite{Chirigati2016} to record: (1) the tree of
processes executed by the pipeline and (2) the list of files read and
written by each process. This information is collected by system call
interceptions, through the \texttt{ptrace} Unix system call, and
stored in a \texttt{SQLite} database. The database stores information
on all the processes which are created by the \texttt{clone()} or
\texttt{fork()} system calls. It also records the files opened by each
process (in read, write or execution mode), including the result files
and the temporary files.

We reconstruct the tree of processes starting from the first process 
created by the pipeline and identifying its child processes as the ones 
that were created through \texttt{clone()} or \texttt{fork()}. We 
create a process graph from this tree by adding edges corresponding to 
file dependencies between processes. A file dependency is defined 
between processes A and B if a file written by A is read by B. 
Figure~\ref{fig:simple_script} shows an example of a process graph 
constructed at this stage.

\medskip
\noindent
\begin{minipage}[]{.5\linewidth}
  \centering
  \includegraphics[scale=0.2,width=.4\linewidth]{images/simple_graph}
%  \captionof{figure}{This is a figure caption.} \label{myfig1}
\end{minipage}
\begin{minipage}[]{.3\linewidth}
\small\captionof{algorithm}{sample script of a process graph}\label{sample_script}
{\scriptsize
\begin{verbatim}
#!/bin/bash
if [ # !=1 ]
then
    echo "usage: 0 <inputimage.nii.gz>"
    exit 1
fi
input_image=$1
bet_output="$(basename ${input_image} .nii.gz)_brain.nii.gz"
bet_output_binarized="$(basename ${input_image} .nii.gz)_brain_bin.nii.gz"

bet ${input_image} ${bet_output} > bet_temp.out
echo "Voxels / Volume in brain mask:"
fslstats ${bet_output} -V
fslmaths ${bet_output} ${bet_output_binarized}
echo "Voxels / Volum in binarized brain mask:"
\rm bet_temp.out
\end{verbatim}
}
\end{minipage}
\captionof{figure}{Process graph constructed from an example brain extraction
	script. Every node in the graph is labeled using (1) a process id 
	created by our reconstruction, (2) the name of the executable run 
	by the process. Processes that read or write temporary files are 
	represented with squares; other ones with circles. Plain edges 
	represent the process tree (\texttt{fork()} or \texttt{clone()} 
	system calls). Dashed edges represent file dependencies: temporary 
	files are in yellow and result files are in green.} 
	\label{fig:simple_script}
\medskip

A different process graph may be produced for every subject in every 
execution condition. Among our dataset, we identified 4 types of 
subjects with different numbers of T1w and T2w images. We verified that 
the process graphs were identical for all subjects of the same type, 
for all execution conditions. The remainder of the 
analysis is done separately for each subject type.

Cycles may be present in the process graph in case a file was written 
by more than one process. We remove such cycles by removing file edges 
between processes A and B when A's process creation timestamp is 
posterior to B's or when A=B. Indeed, such edges cannot happen in 
practice unless A and B were running concurrently, which we assume is 
not the case (we also checked that on the workflow we tested).

In the next step, we classify the processes into three 
categories based on the process graph and the error matrix file 
mentioned in the previous Section:
\begin{enumerate}
\item Processes that read files that do not have errors and write files that do not 
have errors are \emph{transparent} (Figure~\ref{fig:processes}.a).
\item Processes that read files 
that do not have any error but write files that have errors 
\emph{create} errors in the pipeline (Figure~\ref{fig:processes}.b).
\item Processes that read files 
that have errors and write files that do not have errors \emph{remove} 
errors from the pipeline (Figure~\ref{fig:processes}.c).
\item Processes that read files that have errors and write files that also have errors are 
\emph{unknown} (Figure~\ref{fig:processes}.d).
\end{enumerate}

\begin{figure}[H]%\centering
    \begin{subfigure}{0.15\textwidth}
        \includegraphics[scale=0.35]{images/green.png}
        \caption{Transparent}
        \label{fig:green}
    \end{subfigure}
\hfill%\hfil
    \begin{subfigure}{0.15\textwidth}
    \includegraphics[scale=0.35]{images/red.png}
    \caption{Creates error}
    \label{fig:red}
\end{subfigure}
\hfill%\hfil
    \begin{subfigure}{0.15\textwidth}
    \includegraphics[scale=0.35]{images/blue.png}
    \caption{Removes error}
    \label{fig:blue}
\end{subfigure}
\hfill%\hfil
    \begin{subfigure}{0.15\textwidth}
    \includegraphics[scale=0.35]{images/yellow.png}
    \caption{Unknown}
    \label{fig:yellow}
\end{subfigure}
    \caption{Different type of classified process based on the input/output files.
  Dashed edges refer to the file dependencies between processes A and B 
  if a file written by A is read by B. Solid black edges refer to the 
  relationship between parent and child processes.}
    \label{fig:processes}
\end{figure}

\paragraph{Challenge1:} The classification of processes that read or 
write temporary files is uncertain when these files are deleted during 
the execution. To address this issue, we replace every process P that 
writes temporary files with a modification of P that first calls P and 
then backs up all its output files to a read-only directory. This 
replacement is done by modifying the Unix PATH variable to point to a 
directory containing the modified versions of the processes. This 
solution does not cover the temporary files that are removed by P 
itself; this is not a problem since these files do not play any role in 
the subsequent steps of the pipeline, by definition.

\paragraph{Challenge2:} Files written by multiple processes also lead 
to unknown classification labels. To address this issue for a file F 
written by processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
(1) check that processes in \textbf{P} do not write concurrently to F, 
(2) we establish an order on \textbf{P} based on the creation timestamp 
of the processes, (3) we replace ever process $P_{i}$ in \textbf{P} by 
a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
directory. Thus, multiple versions of F are saved and used in the 
analysis.

\paragraph{Challenge3:} Another challenge is to classify processes that 
read files that have errors and write files that also have errors. In 
such situations, it is not possible to determine from the pipeline 
results whether the process created errors, removed some errors, or was 
transparent. We label such processes \emph{unknown}.

To address Challenge 3, we developed an iterative approach that 
consists of the following steps:

@@

\begin{figure}[H]
\centering
  \includegraphics[scale=0.5]{images/iterative_modif}
  \caption{iteration of proposed modification.}
  \label{fig:iterations}
\end{figure}

Figure~\ref{fig:iterations} illustrates our iterative classification 
process for the example in Figure~\ref{fig:graph-example}. At every 
step, processes that created errors are shown in red, processes that 
removed errors are in blue, processes that propagated errors are in 
yellow, and other processes (transparent processes) are in green. As in 
Figure~\ref{fig:graph-example}, plain black edges represent the process 
tree and dashed edges represent file dependencies: green edges 
represent files with no errors, while red edges represent files with 
errors. Temporary files are not represented because they have been 
backed up as previously described.

@@

The process graphs of real pipelines are much larger than the one in 
Figure~\ref{fig:iterations}. To represent such large graphs, we 
summarize them by (1) building subgraphs for every interesting node 
(e.g. red and blue nodes with their children) (2) building connected 
components in the absence of interesting nodes for the remaining nodes 
in the process graph. Each node in the summarized graph represent 
either a shrinked interesting subgraph or connected component and edges 
represent any connection between two shrinked nodes.

\section{HCP Minimal preprocessing Pipelines}

The Human Connectome Project (HCP)~\cite{glasser2013minimal} developed 
a set of pipelines to help extraction of structural, functional or 
diffusion MRI data across a large set of high resolution MR images. 
These pre-processing pipelines create results that are available in 
standard volume and combined surface and volume spaces which makes it 
easier for researchers to compare the images across the neuroimaging 
spectrum.

We focused on HCP structural pre-processing pipelines to (1)quantify 
the effect of OS on HCP pre-processing pipelines (2) identify the 
process(es) in the pipelines that are responsible for such effects.

The structural pre-processing pipelines consists of PreFreeSurfer, 
FreeSurfer and PostFreeSurfer. In this paper, 
we analyzed two pipelines, PreFreeSurfer and FreeSurfer. 
PreFreeSurfer consist of various steps including correction of MR 
gradient nonlinearity distortions, align the T1w and T2w images using 
FSL FLIRT, Align native space to MNI template using ACPC and FSL FLIRT, 
brain extraction usin FSL FLIRT and FNIRT to MNI template, register T2w 
to T1w using FLIRT's BBR, perform a bias field correction, and register 
the subject's native structural space to MNI space.

FreeSurfer steps are downsample the images if it exceeds the resolution 
limit, generation of white matter surface using a segmentation of 
downsampled image, adjust T1w colum intensity and white matter surface 
position based on 0.7mm T1w images, fine tune the T2w to T1w 
registration using FreeSurfer's BBRegister algorithm, accurately place 
the white and pial surface with high-resolution 0.7mm T1w and T2w data, 
these inputs are fed back into recon-all at each stage.

We randomly selected 10 subjects from the HCP data release S500 which 
are described in table~\ref{table:data}.

\begin{table}[H]\hspace*{-1cm}
\centering
\begin{threeparttable}
\caption{An overview of the HCP data (Humman Connectome DB).}

\begin{tabular}{@{}llllll@{}}
\toprule
Subject & Release & Acquisition & Gender & Age      & Subj\_Type   \\ \midrule
103515  & Q1      & Q02         & F      & 26-30    & type1         \\
105216  & Q3      & Q03         & M      & 26-30    & type4         \\
103414  & Q2      & Q02         & F      & 22-25    & type1         \\
125525  & Q1      & Q01         & F      & 31-35    & type1         \\
142828  & Q1      & Q01         & M      & 31-35    & type1         \\
129533  & Q3      & Q04         & F      & 31-35    & type1         \\
103818  & Q1      & Q01         & F      & 31-35    & type3         \\
133928  & Q2      & Q03         & M      & 26-30    & type3         \\
148032  & Q3      & Q03         & F      & 31-35    & type2         \\
139637  & Q2      & Q03         & F      & 31-35    & type1         \\ \bottomrule
\end{tabular}
\begin{tablenotes}
      \small
      \item *Subj\_Type refers to the graph topology type of the subjects,
      we will discuss further on the next sections.
\end{tablenotes}
\end{threeparttable}
\label{table:data}
\end{table}

\section{Results}

We executed the pipelines using Docker containers to simplify the 
deployment of different operating system versions on execution 
platforms. The Docker images were built for the HCP pre-processing 
pipelines v3.19.0 (PreFreeSurfer and FreeSurfer) in 
CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for 
reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}. 
We collected the provenance trace (tree of executed processes and files 
accessed in read or write mode) for each type of 
subjects(Table~\ref{table:data}) using system-call interception as 
provided by the \reprozip tool~\cite{Chirigati2016}.

Two types of errors can occur in the subjects due to the 
errors in the operating systems. One is inter-OS error caused by the 
operating system library updates and the other type, intra-OS errors 
occurs as a result of the pseudo-random processes used in the 
pipelines. An example of a pseudo-random process function is, a random 
number generator that would get initialized using a seed state. the 
proposed method can be used to identify both kind of errors. The files 
that are common to all the subjects only are taken into consideration 
for comparison. The first step is identification of files with 
differences in their checksums. This is identified using the checksums 
that are recorded after the processing. Intra-OS errors are identified 
using the run-number added as the suffix for the conditions. For 
example, the two batches of subjects processed under the same condition 
(CentOS6) are stored as run-1 and run-2. The files belonging to the 
subjects stored under the above mentioned conditions are treated as 
intra-OS runs.

\subsection{Inter-OS binary differences}

In the previous study~\cite{Scaria2017}, we showed that pre-processing 
pipelines of the Human Connectome Project~\cite{Glasser2013} were 
sensitive to operating system variation so that Figure 
\ref{fig:tissue_class} illustrates the binarized errors after brain 
tissue segmentation process in FreeSurfer. In addition, 
Figure~\ref{fig:fnirt_result} shows the same errors occurred in 
non-linear registration process between CentOS6 and CentOS7 in PreFreeSurfer.

\begin{figure}[H]
%  \includegraphics{brain\_classification}
\centering
  \includegraphics[scale=0.18]{images/brain_classification.png} 
  \caption{Binarized error between brain segmentation results from 
  FreeSurfer, subject 105216 (CentOS6 vs. CentOS7)~\cite{Scaria2017}.
    } 
  \label{fig:tissue_class}
\end{figure}

\begin{figure}[H]
\centering
  \includegraphics[scale=0.6]{images/fnirt_result.png} 
  \caption{Absolute errors between FNIRT results from PreFreeSurfer 
  (T2w\_acpc\_to\_MNI\_nonli.nii.gz), subject 104820 (CentOS6 vs. 
  CentOS7)~\cite{Scaria2017}. } 
  \label{fig:fnirt_result}
\end{figure}

\subsection{Pipeline analysis}

We identified processes that introduce errors in both PreFreeSurfer and 
FreeSurfer pipelines. The result of our experiment is described in the 
following paragraphs.

\subsubsection{PreFreeSurfer pipeline}

Among the 117 data files produced by 
PreFreeSurfer, 21 did not have any error for any subject, 92 had errors 
for all subjects and 4 had errors for 3 subjects only.

Figure \ref{fig:summarized-graph} shows the annotated provenance graph 
of the PreFreeSurfer pipeline executed on CentOS6 and CentOS7. Each 
node in the graph represent an executed process in the pipeline. 
Processes that created errors are shown in red, processes that removed 
errors are in blue, and other processes are in green.  Squares denote 
processes for which the classification is uncertain, due to temporary 
files that were removed during the execution. Black edges link 
sub-processes to their parents while dashed edges denote file 
dependencies between processes (green edges: files with no errors; red 
edges: files with errors; yellow edges: temporary files).

>>>>

% GigaScience template
\documentclass[a4paper,num-refs]{oup-contemporary}

\journal{gigascience}

%%%% Packages %%%%
\usepackage{siunitx}
\usepackage{minted} % Used for JSON highlighting
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}
\usepackage{booktabs}

%%%%%%

\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{makecell}
\usepackage[flushleft]{threeparttable}

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip}

\title{NRM-tool (Numerical Reproducibility Measurement): 
a debugging tool to identify numerical errors 
and assess reproducibility of pipelines}

@@

\author[1]{Ali Salari}
\author[1]{Lalet Scaria}
\author[2,3]{Gregory Kiar}
\author[2]{Lindsay Lewis}
\author[1]{Tristan Glatard}
\author[2,3]{Alan C. Evans}

\affil[1]{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}
\affil[2]{McGill University, Montreal, Canada}
\affil[3]{Montreal Neurological Institute, Montreal, Canada}

@@

\begin{abstract}
Many experiments show that computational analysis results are still not 
completely reproducible. In particular, computational environments 
including different operating systems, hardware categories, and 
software versions are known to have effects on the results produced by 
analysis pipelines. These effects are presumably due to the creation, 
propagation and amplification of small numerical errors across the 
pipelines. 
Although new techniques such as virtualization , version 
control systems , and provenance management tools provide more reliable 
environment and significantly improve the reproducibility of scientific 
findings. 
However, the precise causes of such instabilities and the path along 
which they propagate in the pipelines are unclear.  We present a 
technique to identify the processes in the pipeline that create 
numerical errors along the execution, and we apply this technique to 
the HCP structural pre-processing pipelines.

\end{abstract}

\begin{keywords}
Reproducibility; Numerical Instability; Neuroimaging.
\end{keywords}

@@

A reproducible study provides a context in which one is able to get 
results that are consistent with the original 
work~\cite{plesser2018reproducibility}. Research findings are expected 
to be reproducible so that their authenticity and reliability can be 
evaluated. 
There are different definitions for the term reproducibility, we follow 
Peng's definition~\cite{peng2011reproducible} in which reproducibility 
is defined as the ability to regenerate the same results as the 
original findings when the experiment is reanalyzed given exactly the 
same analytic methods, software package, parameters, and data.

In addition, numerical reproducibility is defined as the ability to 
regenerate bit for bit identical results from multiple 
runs~\cite{hill2017numerical}. Two files will be considered numerically 
reproducible if they have identical binary contents. Binary comparison 
is calculated by comparing checksums. It must be pointed that a 
computation might be reproducible based on Peng's definition, but not 
be numerically reproducible. For instance, small numerical errors 
created during the pipeline execution may hamper numerical 
reproducibility, but be negligible in the final results.

Recently, scientists began to realize that the results of many 
scientific experiments were not reproducible. Results show that the 
variety of computational infrastructures including workstation types, 
parallelization methods, operating systems, and analysis packages are 
known to influence reproducibility because of the creation of small 
numerical errors~\cite{Gronenschild2012, diethelm2012limits, 
Glatard2015, bowring2018exploring}. The propagation and amplification 
of these tiny errors by analysis pipelines may cause reproducibility 
issues. 
In this case, the analysis pipelines are said to be numerically 
unstable. Numerical instability is a characteristic of the pipelines 
which amplify small numerical errors and then hamper the 
reproducibility of the analyses depending on the length of the pipeline 
and magnitude of the errors. In many cases, numerical instability is an 
important issue for reproducibility.

In particular, the effect of operating system is quantified 
in~\cite{Glatard2015, Gronenschild2012} more specifically on some of 
the main neuroimaging pipelines including FSL, CIVET and Freesurfer 
pipelines. The results of comparisons indicate a significant errors 
between analysis results of the same experiment on different operating 
system. It should be noted that \emph{error} is defined as the binary 
differences between results of two analyses in this paper, which is 
calculated by the checksum method. 
The operating system is not the only part of the computational 
environment that may hamper reproducibility. Many other computational 
factors including hardware configuration, software version, and 
parallelization techniques are reported as the main influential 
computing infrastructure on reproducibility.

There are different solutions to make reproducible analysis including 
containerization techniques to encapsulate software/hardware 
dependencies, provenance capturing tools, and version control systems. 
However, a comprehensive solution requires to fix the numerical 
instabilities. For this purpose, we introduce a Numerical 
Reproducibility Measurement (NRM) tool to identify the processes in 
pipeline that create numerical errors across different runs. In this 
paper, the execution results of pipelines across different operating 
systems are considered to find out the cause of the irreproducibility 
using an interposition techniques.

The reminder of this paper describe the NRM-tool in different parts as 
the way of capturing provenance information and their representation, 
the influence of different type of subjects, and how to classify 
processes and characterize errors. We describe the HCP preprocessing 
pipelines as a popular neuroimaging pipelines to test and apply the 
NRM-tool on. Finally, we report the experimental results, the processes 
that are responsible for the errors in the pipelines.

\section{Tool description}

We present a Numerical Reproducibility Measurement tool (NRM-tool) to 
measure numerical reproducibility of the pipelines across operating 
systems automatically (see Figure~\ref{fig:overview}). We build the 
containerized version of pipelines for each operating system using 
docker technology which encapsulated whole the pipeline dependencies. 
The docker images consists of an specific computing environment such as 
operating system version, hardware configuration, software version, 
etc. Furthermore, to simplify computational reproducibility of the 
analysis, we execute docker images using Boutiques framework based on 
the command-line descriptions~\cite{glatard2017boutiques}. The 
Boutiques descriptor specifies the parameters and implementation of the 
experiment, and the invocation schema describes the parameter values 
such as input/output data path.

NRM-tool produces the error JSON schema in which recognizes output files 
with errors using the checksum comparison of the results. 
We consider that a file contains errors if 
its binary comparison is different, otherwise file has no error. Also, 
we use \reprozip tool to capture the provenance information. 
Subsequently, using a iterative approach, we classify processes into 
different categories based on the state of read/write files 
with/without error. NRM-tool identifies relevant processes in the 
pipeline creating, propagating, or removing errors. We will explain 
each steps in detail in the following sections.

\begin{figure}
\centering
  \includegraphics[width=\columnwidth]{images/overview.png}
  \caption{An overview of the proposed technique.}
  \label{fig:overview}
\end{figure}

\subsection{Provenance capture and representation}

We use the \reprozip tool~\cite{Chirigati2016} to record: (1) 
the tree of processes executed by the pipeline and (2) the list of 
files read and written by each process. This information is collected 
by system call interceptions, through the \texttt{ptrace} Unix system 
call, and stored in a \texttt{SQLite} database. The database stores 
information on all the processes which are created by the 
\texttt{clone()} or \texttt{fork()} system calls. It also records the 
files opened by each process (in read, write or execution mode), 
including the result files and the temporary files.

NRM-tool reconstructs the tree of processes starting from the first process 
created by the pipeline and identifying its child processes as the ones 
that were created through \texttt{clone()} or \texttt{fork()}. It 
creates a process graph from this tree by adding edges corresponding to 
file dependencies between processes. A file dependency is defined 
between processes A and B if a file written by A is read by B. 
Figure~\ref{fig:simple_script} shows an example of a process graph 
constructed for the simple bash script illustrated in 
Algorithm~\ref{algo:sample-script} from the brain extraction procedure.

\begin{algorithm}[h!]
\caption{Sample script from brain extraction process}
\label{algo:sample-script}
\begin{verbatim}
#!/bin/bash
if [ # !=1 ]
then
    echo "usage: 0 <inputimage.nii.gz>"
    exit 1
fi
input_image=$1
bet_output="$(basename ${input_image} 
                       .nii.gz)_brain.nii.gz"
bet_output_binarized="$(basename ${input_image} 
                       .nii.gz)_brain_bin.nii.gz"

bet ${input_image} ${bet_output} > bet_temp.out
echo "Voxels / Volume in brain mask:"
fslstats ${bet_output} -V
fslmaths ${bet_output} ${bet_output_binarized}
echo "Voxels / Volum in binarized brain mask:"
\rm bet_temp.out
\end{verbatim}
\end{algorithm}

\begin{figure}
\centering
  \includegraphics[scale=0.3]{images/simple_graph}
  \caption{Process graph constructed from an example brain extraction
	script. Every node in the graph is labeled using (1) a process id 
	created by our reconstruction, (2) the name of the executable run 
	by the process. Processes that read or write temporary files are 
	represented with squares; other ones with circles. Plain edges 
	represent the process tree (\texttt{fork()} or \texttt{clone()} 
	system calls). Dashed edges represent file dependencies: temporary 
	files are in yellow and result files are in green.}
  \label{fig:simple_script}
\end{figure}

\subsection{Subject type}

Depending on the variety of computing environments, the produced 
process graph can be different. In particular, different analysis 
software package produce different process graphs for the similar 
experiments because of following different methods. We used a fixed 
analysis software package in different conditions, however, we found 
that different subjects may produce a different process graph. Among 
our dataset, we identified 4 types of subjects with different numbers 
of T1w and T2w images.

In addition, we verified that the process graphs were identical for all 
subjects of the same type in different versions of CentOS operating 
system. The remainder of the analysis is done separately for each 
subject type.

\subsection{Graph analysis}

NRM-tool classifies the processes into four 
categories based on the process graph and the error JSON file 
mentioned in the previous sections:
\begin{enumerate}
\item Processes that read files that do not have errors and write files 
that do not have errors are \emph{transparent} (Figure~\ref{fig:processes}.a).
\item Processes that read files 
that do not have any error but write files that have errors 
\emph{create} errors in the pipeline (Figure~\ref{fig:processes}.b).
\item Processes that read files 
that have errors and write files that do not have errors \emph{remove} 
errors from the pipeline (Figure~\ref{fig:processes}.c).
\item Processes that read files that have errors and write files that also have errors are 
\emph{unknown} (Figure~\ref{fig:processes}.d).
\end{enumerate}

\begin{figure}%\centering
\centering
    \begin{subfigure}{0.4\linewidth}
        \includegraphics[scale=0.34]{images/green.png}
        \caption{Transparent}
        \label{fig:green}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.4\linewidth}
    \includegraphics[scale=0.34]{images/red.png}
    \caption{Creates error}
    \label{fig:red}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.4\linewidth}
    \includegraphics[scale=0.34]{images/blue.png}
    \caption{Removes error}
    \label{fig:blue}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.4\linewidth}
    \includegraphics[scale=0.34]{images/yellow.png}
    \caption{Unknown}
    \label{fig:yellow}
\end{subfigure}
    \caption{Different type of classified process based on the input/output files.
  Dashed edges refer to the file dependencies between processes A and B 
  if a file written by A is read by B. Solid black edges refer to the 
  relationship between parent and child processes.}
    \label{fig:processes}
\end{figure}

There are some issues that should be addressed to classify process 
properly, including capturing temporary files that have been removed 
during the execution, capturing files written by multiple processes, 
and unknown process.

\subsubsection{Capture temporary files} 
The classification of processes that read or 
write temporary files is uncertain when these files are deleted during 
the execution. To address this issue, we replace every process P that 
writes temporary files with a modification of P that first calls P and 
then backs up all its output files to a read-only directory. This 
replacement is done by modifying the Unix PATH variable to point to a 
directory containing the modified versions of the processes. This 
solution does not cover the temporary files that are removed by P 
itself; this is not a problem since these files do not play any role in 
the subsequent steps of the pipeline, by definition.

\subsubsection{Capture multiple write files} 
Files written by multiple processes also lead 
to unknown classification labels. To address this issue for a file F 
written by processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
(1) check that processes in \textbf{P} do not write concurrently to F, 
(2) we establish an order on \textbf{P} based on the creation timestamp 
of the processes, (3) we replace ever process $P_{i}$ in \textbf{P} by 
a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
directory. Thus, multiple versions of F are saved and used in the 
analysis.

Furthermore, cycles may be present in the process graph in case a file 
was written by more than one process. We remove such cycles by removing 
file edges between processes A and B when A's process creation 
timestamp is posterior to B's or when A=B. Indeed, such edges cannot 
happen in practice unless A and B were running concurrently, which we 
assume is not the case (we also checked that on the workflow we 
tested).

\subsubsection{Unknown process} 
Another challenge is to classify processes that 
read files that have errors and write files that also have errors. In 
such situations, it is not possible to determine from the pipeline 
results whether the process created errors, removed some errors, or was 
transparent. We label such processes \emph{unknown}.
To address this issue, we developed an iterative approach that 
consists of the following steps:

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{images/iterative_modif}
  \caption{iteration of proposed modification.}
  \label{fig:iterations}
\end{figure}

Figure~\ref{fig:iterations} illustrates our iterative classification 
process for the example in Figure~\ref{fig:simple_script}. At every 
step, processes that created errors are shown in red, processes that 
removed errors are in blue, processes that propagated errors are in 
yellow, and other processes (transparent processes) are in green. As in 
Figure~\ref{fig:simple_script}, plain black edges represent the process 
tree and dashed edges represent file dependencies: green edges 
represent files with no errors, while red edges represent files with 
errors. Temporary files are not represented because they have been 
backed up as previously described.

@@

\section{Experiments}
\subsection{HCP pipelines and datasets}
\subsection{Results}
